\section{Abstract}

Car sharing is a popular means of transport in smart cities, with the free floating paradigm which lets the customers autonomously pick and drop available cars. This freedom impose several challenges and trade-off in the design of an electrical version of this type of car sharing. 

In this paper we study the design of an electric free floating car sharing system for four cities by leveraging actual rental traces. We analyze via accurate simulations the impact of i) the charging station placement, and ii) return policies.
Our traces composed by millions of rentals in different cities world wide located, captures the non-stationary and highly dynamic nature of usage patterns of customers.

Considering charging station placement, we demonstrate how it is better to place charging stations within popular parking area (e.g., downtown), even if parking duration is short. We show how by smart placing and handful number of charging stations (below 10\% of city zones) the system can self-sustain itself. 
These results are obtained also thanks to customers' willingness to collaborate by voluntarily returning the car to a nearby charging station.

Our results help car sharing provider comparing possible alternative design solution, i.e.,  the adoption of simple relocation policies that would move cars that need a charge only, a promising solution to limit discomfort for customers due to rerouting. The same could be achieved by considering giving incentives to customers. We offer all data and tools used in this paper to foster further studies in these directions.

